paper1_figures

### 2023-10-18 paper1_markdown.Rmd
18.10 mac

author: 
  - name          : "Pedrazzini Sabina"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Bern, Centre for Development and Environment (CDE), Mittelstrasse 43, CH-3012 Bern"
    email         : "sabina.pedrazzini@unibe.ch"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"

@@

In conclusion, social status impacts cooperation, but it seems to have multiple effects. Moreover, as mentioned in the previous section, @Sell2009 affirmed that, when investigating gender differences in cooperation, it is impossible to disentangle the effect of social status. Similarly, @Aksoy2019 demonstrated that the lower level of cooperation between members of different social classes is mainly due to actors' identification with their classes. This shows that social identity and social status are deeply connected, and that therefore it is difficult to separate their effects on cooperation.

@@

First, according to the analyzed papers, trust has a positive effect on cooperation. In fact, since cooperation involves the risk that others will defect, trusting others is essential [@Henry2011]. Moreover, according to the principle of reciprocity, people cooperate more when facing cooperative partners [@Fehr2022; @VanLange1999]. However, in real-life situation it is not always possible to know exactly others' contributions, and therefore trust (i.e. beliefs about how much others cooperate) become essential to decide how to behave [@Vuolevi2012]. Moreover, trust among participants of a common-pool setting is at core of the framework of @Poteete2010, aimed at explaining why commons can (not) be successful. These evidences fit well with the more general result showing that dispositional trust is a predictor of pro-social behaviour [@Wu2020]. Finally, this link has a biological grounding: being trusted increases the level of oxytocin (a hormone that helps creating attachment between individuals), and this, in turn, improve the cooperative behaviour toward the person who first showed trust [@Campbell2010].

@@

Concerning gender effect on cooperation, the existing literature about gender stereotypes would predict that women cooperate more than men. In fact, according to the stereotypes, women are more communal and caring, while men are more agentic and performance-oriented [@Eagly2012], and this lead to the perception that women's intentions are more benevolent than those of men [@Ellemers2017]. In line with these predictions, @Westermann2005 found that, managing a natural resource, groups where women were present displayed more solidarity and collaboration comparing to all-men groups. Despite these evidences, this review found that there is no general gender effect on cooperation and that a lot of moderator are at stake. Moreover, gender is always embedded with social status dynamics [@Sell2009], and it is therefore difficult to separate the two effects. The analyzed studies about social status and cooperation showed that usually members of low-status groups cooperated more and showed more solidarity with others low-status members than high-status. Although the effect of social status on cooperation has not been extensively studied, these results seem to be consistent with the existing literature on social status. For instance, members of high-status groups are more individualistic, because their group belonging does not reduce individual characteristics as it does with low-status [@Lorenzi-Cioldi2002]. Moreover, high-status individuals display more system justification than low status, i.e. they see the social system as fair and secure [@Caricati2012], and higher levels of Social Dominance Orientation [@Vargas-Salfate2018], that is the personal preference for inequality between social groups and that is negatively correlated with concern about others and altruism [@Pratto1994]. Finally, although the differences between social statuses, people tend to cooperate more with individuals of their same social class than with others, and this effect is manly due to group identification [@Aksoy2019]. Social status is therefore deeply connected to the last of the identified individual factors, group identification.

The reviewed results showed that the more individuals identify with their group, the more they are likely to act for the group's good. In fact, people's identity is also composed by their group affiliations, and the stronger they identify a social group, the more important this group becomes for their identity and the more their personal interests aligns with group's ones [@Tajfel1974]. Also, within a social group, individuals perceive themselves are more similar to other members [@Turner1984], and therefore they are more susceptible to adopt collective goals and behaviours [@Fritsche2018]. Moreover, strongly identified people are less likely to leave a social group in case of failure or difficulties [@Ellemers1997]. Thus, in order for a social group to be successful, its members must strongly identify with it. In fact, in-group identification is the main predictor of collective action in the Social Identity Model of Collective Action [SIMCA, @VanZomeren2008]. Moreover, @Fritsche2018 apply this social identity approach to environmentalism with the Social Identity Model of Pro-Environmental Action (SIMPEA). According to its authors, the climate crisis has the structure of common good dilemmas, because environmental goals can be attained only if everybody cooperates. Group identification is once again the main predictor of such collective actions.

@@

Until now, the identified factors has been discussed separately. However,

@@

Finally, all of the group and individual factors identified are deeply interconnected.

-   For example, women level of cooperation is predicted by their trust in their interaction partner, while men's trust and cooperation are not linked [@Irwin2015]. Moreover, women's lower cooperation is explained by the fact that they expected their partner to contribute less to the common good [@Dorrough2019].

@@

@Janssen2015: In contrast to lab experiments, studies find that cultural factors, experiences with specific institutional arrangements and ecological conditions can affect the way participants make decisions in those experiments. This demonstrates that effectiveness of interventions is context specific.

However, it is not without limitations, thus further researches are necessary to reach a deeper understanding of the topic. First, most of the studies included had a laboratory social dilemma setting, who do not allow to exactly reproduce the dynamics present in a commons [@Cole2010]. Therefore, this papers offers a good basis about elements influencing cooperation in commons, but future studies should confirm this results in real-life settings. Moreover, this review is not informative about the relative importance of each factor. To shed light on this topic, quantitative studies are needed.

-   @Henry2011: Such a theory must build on research in less complex contexts, but must also give attention to the added effects of evolving networks and information-sharing amongst those making decisions relevant to sustainability and global environmental change (their study was on trust, but it can be applied to this review in general)

>>>>

author: 
  - name          : "Pedrazzini Sabina"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Bern, Centre for Development and Environment (CDE), Mittelstrasse 43, CH-3012 Bern"
    email         : "sabina.pedrazzini@unibe.ch"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
     # - "Conceptualization"
     # - "Writing - Original Draft Preparation"
     # - "Writing - Review & Editing"

@@

In conclusion, social status impacts cooperation, but it seems to have multiple effects. Moreover, similarly to what @Sell2009 affirmed about the effects of gender and social status, @Aksoy2019 demonstrated that the lower level of cooperation between members of different social classes is mainly due to actors' identification with their classes. This shows that social identity and social status are deeply connected, and that therefore it is difficult to separate their effects on cooperation.

@@

First, according to the analyzed papers, trust has a positive effect on cooperation. In fact, since cooperation involves the risk that others will defect, trusting others is essential [@Henry2011]. Moreover, according to the principle of reciprocity, people cooperate more when facing cooperative partners [@Fehr2022; @VanLange1999]. However, in real-life situation it is not always possible to know exactly others' contributions, and therefore trust (i.e. beliefs about how much others cooperate) becomes essential to decide how to behave [@Vuolevi2012]. Moreover, trust among participants of a common-pool setting is at core of the framework of @Poteete2010, aimed at explaining why commons can (not) be successful. These evidences fit well with the more general result showing that dispositional trust is a predictor of pro-social behaviour [@Wu2020]. Finally, this link has a biological grounding: being trusted increases the level of oxytocin (a hormone that helps creating attachment between individuals), and this, in turn, improve the cooperative behaviour toward the person who first showed trust [@Campbell2010].

@@

Concerning gender effect on cooperation, the existing literature about gender stereotypes would predict that women cooperate more than men. In fact, according to the stereotypes, women are more communal and caring, while men are more agentic and performance-oriented [@Eagly2012], and this lead to the perception that women's intentions are more benevolent than those of men [@Ellemers2017]. In line with these predictions, @Westermann2005 found that, managing a natural resource, groups where women were present displayed more solidarity and collaboration comparing to all-men groups. Despite these evidences, this review found that there is no general gender effect on cooperation and that a lot of moderator are at stake. Moreover, gender is always embedded with social status dynamics [@Sell2009], and it is therefore difficult to separate the two effects. The analyzed studies about social status and cooperation showed that usually members of low-status groups cooperated more and showed more solidarity with others low-status members than high-status. Although the effect of social status on cooperation has not been extensively studied, these results seem to be consistent with the existing literature on social status. For instance, members of high-status groups are more individualistic, because their group belonging does not reduce individual characteristics as it does with low-status [@Lorenzi-Cioldi2002]. Moreover, high-status individuals display more system justification than low status [@Caricati2012], and higher levels of Social Dominance Orientation [@Vargas-Salfate2018]. Therefore, they see the social system as fair and secure and they have a preference for inequality between social groups. This is negatively correlated with concern about others and altruism [@Pratto1994]. Finally, although the differences between social statuses, people tend to cooperate more with individuals of their same social class than with others, and this effect is manly due to group identification [@Aksoy2019]. Social status is therefore deeply connected to the last of the identified individual factors, group identification.

The reviewed results showed that the more individuals identify with their group, the more they are likely to act for the group's good. In fact, people's identity is also composed by their group affiliations, and the stronger they identify with a social group, the more important this group becomes for their identity and the more their personal interests aligns with group's ones [@Tajfel1974]. Also, within a social group, individuals perceive themselves as more similar to other members [@Turner1984], and therefore they are more susceptible to adopt collective goals and behaviours [@Fritsche2018]. Moreover, strongly identified people are less likely to leave a social group in case of failure or difficulties [@Ellemers1997]. Thus, in order for a social group to be successful, its members must strongly identify with it. In fact, in-group identification is the main predictor of collective action in the Social Identity Model of Collective Action [SIMCA, @VanZomeren2008]. Moreover, @Fritsche2018 apply this social identity approach to environmentalism with the Social Identity Model of Pro-Environmental Action (SIMPEA). According to its authors, the climate crisis has the structure of common good dilemmas, because environmental goals can be attained only if everybody cooperates. Group identification is once again the main predictor of such collective actions.

@@

Until now, the identified factors has been discussed separately. However, their impacts are interconnected, as already anticipated in some reviewed results about moderation effects. For instance, incentives are more useful to improve cooperation of low-, rather than high-identified, group members [@VanVugt2009]; trust is a better predictor of cooperation in women than in men [@Irwin2015; @Dorrough2019]. These connections are also confirmed in the litterature.

@@

@@

However, it is not without limitations, thus further researches are necessary to reach a deeper understanding of the topic. First, most of the studies included had a laboratory social dilemma setting, who are simplier than real-life situations and do not allow to exactly reproduce the dynamics present in a commons [@Henry2011; @Cole2010]. In fact, some elements such as cultural factors and specific ecological conditions affect the decision making process, and cannot be reproducted in lab experiments [@Janssen2015]. Therefore, this papers offers a good basis about elements influencing cooperation in commons, but future studies should confirm this results in real-life settings. Moreover, this review is not informative about the relative importance of each factor. To shed light on this topic, quantitative studies are needed.
